### 2015-07-07 MSH.tex
Update

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\usepackage{pdflscape}
\modulolinenumbers[5]

\journal{Geochemica and Cosmochemica Acta}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Hydrous transition zone melting  
%\title{High-pressure water-rich melting in the system MgO-SiO$_2$-H$_2$O: Insights into hydrous transition zone melting}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author{R. Myhill, D. J. Frost}
\address{Bayerisches Geoinstitut, Universit\"{a}t Bayreuth, Universit\"{a}tsstrasse 30, 95447 Bayreuth, Germany}
\cortext[mycorrespondingauthor]{Corresponding author: R. Myhill}
\ead{myhill.bob@gmail.com}

\author{D. Novella}
\address{Laboratoire Magmas et Volcans, Universit\'{e} Blaise Pascal, 5 Rue Kessler, 63038 Clermond-Ferrand, France}

%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}

%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}

%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}

In this study, we conducted high pressure multianvil experiments at 13 GPa between 1200 and 1900 $^{\circ}$C to investigate the liquidus in the system MgO-SiO$_2$-H$_2$O. Water-rich starting compositions were obtained by using platinic acid (H$_2$Pt(OH)$_6$) as a novel water source. As MgO:SiO$_2$ ratios decrease, the lquidus curve develops an increasingly pronounced concave-down topology. At slab temperatures, the forsterite- and enstatite-water liquidii have similar water contents. Preliminary experiments on the SiO$_2$-H$_2$O binary reveal a pronounced melting point depression up to about 20 mol\% H$_2$O. This melting point depression exceeds estimates from published models, and requires either significant non-ideality between species in the melt, or distinct differences in the ease of protonation of bridging oxygens. We discuss the implications of this work on the migration of deep-seated slab fluids, and the partitioning of water between nominally anhydrous phases and melts.
\end{abstract}

\begin{keyword}
high pressure \sep mantle \sep hydrous melting \sep water \sep liquidus
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}
Subduction zones are primary vectors for water transport into the deep Earth. This water is progressively lost as fluids and then melts as the surrounding mantle warms the slab. The most prominent surface expression of this melting is the presence of arc volcanoes about 100 km above the slab surface, but hydrous minerals within the crust should allow some water to be transported to 250 km, and peridotite hydration should allow water to penetrate into the lowermost mantle transition zone, providing it remains low enough in temperature. A startling confirmation of such deep water is the discovery of water-saturated ringwoodite in a kimberlitic diamond from Juina, Brazil.

Several key questions present themselves. When water-rich melts are created in subduction zones, what are their compositions? How do their compositions and resulting physical properties (density, viscosity, melt-solid dihedral angle) change as they migrate? Where will they migrate? Their migration paths will have major implications for figuring out the water budget of the Earth - if they are trapped close to or within the slab, hydrating the whole of the mantle transition zone will be much less likely.

There is another reason to be interested in the compositions of hydrous melts. The thermal stability of dense hydrous magnesium silicate (DHMS) phases and the partitioning of water between silicate melt and nominally anhydrous minerals (NAMS) control water transport into the deep mantle. At shallow depths, the fluids released during dehydration of these phases contain only a few percent of dissolved solutes. As such, it is reasonable to simplify thermodynamic calculations by approximating the fluid product of dehydration as pure water. As pressure increases, several processes ensure that this is no longer the case. Firstly, the second critical endpoint of peridotitic and eclogitic systems occur at <10 GPa, after which there is no longer a distinction between melt and fluid, and melt compositions smoothly migrate towards lower water contents with increasing temperature. Secondly, the thermal stability of DHMS phases increases. As a result, the final breakdown of DHMS phases at mantle transition zone pressures occurs when MgO-SiO$_2$-H$_2$O melts have H$_2$O activities significantly lower than one, which will also strongly influence the water content of NAMS. Accurate thermodynamic estimates of water storage capacities in mantle rocks thus require models for both the silicate phases and the coexisting melts.

% There has been some controversy over whether slabs transport water beyond about 300 km depth. Certainly most of the ca. 7 km thick mafic crust is probably too warm to sustain hydrous minerals. However, hydrated ultramafics can potentially carry water into the lower mantle. Supporting evidence for an ultradeep part of the water cycle exists in the form of hydrous ringwoodite in diamond, which apparently formed under cold conditions.

% The compositions of melts formed by the breakdown of hydrous phases is important for a full understanding of the deep water system. The thermodynamics of melting in water-rich systems cannot be determined without good estimates of melt compositions. Furthermore, the thermal stability of hydrous phases cannot be predicted from their properties without also having an activity-composition model for the melt phase. More generally, the density of hydrous melts will be highly composition dependent, impacting on their ability to rise through the mantle. Finally, reactions between migrating melts and the surrounding solid will be controlled by the \emph{P-T-X} dependence on melt composition.

Constraining melt compositions remains a major challenge in high-pressure experimental petrology. The primary problem is that hydrous melts are unquenchable, which means that their compositions must be estimated indirectly. Compositions of partial hydrous melts have been estimated using mass balance, diamond traps \citep{MSUP2007}, and the Mg:Si ratio has been estimated by defocusing the beam of an electron probe microanalyser (EPMA) \citep[e.g.][]{YII2004}. Unfortunately, all three of these techniques are accompanied by significant uncertainties, especially in terms of the water content. An alternative technique is to constrain the liquidus temperature at a given composition, thus circumventing the need to estimate water concentrations in the melt after the experiment (when much of the water has been lost). The problem with this technique is that at ambient pressures, hydrous phases are limited in their maximum water content. In the MgO-SiO$_2$ system, the lack of a thermally stable stoichiometric SiO$_2$.H$_2$O compound means that brucite has been used as the water source. This limits water concentrations in model peridotitic systems (Mg$\sim$1.4 Si) to $\sim$18 wt\% H$_2$O. Thus, studies either extrapolate high temperature liquidus curves to high water contents or add water in liquid form. Extrapolation is prone to large errors, and accurately adding liquid water to the small volumes of high pressure capsules is essentially impossible.

In this study, we use high purity platinic acid, hexahydroxyplatinate(IV) (H$_2$Pt(OH)$_6$) as a novel water source. Platinic acid is a pale yellow compound which is stable to about 130 $^{\circ}$C \citep{Nagano2002}. It breaks down at higher temperatures, releasing four molecules of H$_2$O and forming platinum (IV) oxide PtO$_2$. Its relatively high stability, high water content and the inert nature of the breakdown product (in a system where redox reactions are absent) make it an excellent water source for high pressure hydrous melting experiments.

\section{Experimental and analytical methods}
Starting compositions were created from a mixture of high purity brucite (Mg(OH)$_2$, SiO$_2$ and platinic acid (H$_2$Pt(OH)$_6$). SiO$_2$ was dried overnight at 1000 $^{\circ}$C, while brucite was heated to 250 $^{\circ}$C overnight. Both powders were stored in a desiccating oven at 130 $^{\circ}$C. Platinic acid is hygroscopic, so was stored in a vacuum desiccator. Powders were weighed and ground dry in an agate mortar for 30 minutes, using a mask, goggles and gloves to avoid physical contact with the platinic acid. Starting compositions were stored in glass vials in a vacuum desiccator. Compositions are listed in Table \ref{table:compositions}.

\begin{table}[ht!]
\caption{Starting compositions}
\label{table:compositions}
\begin{tabular}{l|lll|lll}
 & \multicolumn{3}{c}{Oxide proportions (mol\%)} & \multicolumn{3}{|c}{Compound fractions (mol/mol)} \\
\hline
 & \multicolumn{1}{c}{MgO} & \multicolumn{1}{c}{SiO$_2$} & \multicolumn{1}{c}{H$_2$O} & \multicolumn{1}{|c}{Mg(OH)$_2$} & \multicolumn{1}{c}{SiO$_2$} & \multicolumn{1}{c}{H$_2$Pt(OH)$_6$} \\
\hline
br 5.0 & 50.000 & 0.000 & 50.000 & 1.000 & 0.000 & 0.000 \\
br 5.5 & 45.000 & 0.000 & 55.000 & 0.947 & 0.000 & 0.053 \\
br 6.0 & 40.000 & 0.000 & 60.000 & 0.889 & 0.000 & 0.111 \\
br 6.5 & 35.000 & 0.000 & 65.000 & 0.824 & 0.000 & 0.176 \\
br 7.0 & 30.000 & 0.000 & 70.000 & 0.750 & 0.000 & 0.250 \\
en 6.0 & 30.000 & 30.000 & 40.000 & 0.480 & 0.480 & 0.040 \\
en 7.0 & 26.667 & 26.667 & 46.667 & 0.457 & 0.457 & 0.086 \\
en 8.0 & 23.333 & 23.333 & 53.333 & 0.431 & 0.431 & 0.138 \\
en 9.0 & 20.000 & 20.000 & 60.000 & 0.400 & 0.400 & 0.200 \\
fo 11.5 & 36.000 & 18.000 & 46.000 & 0.637 & 0.319 & 0.044 \\
fo 13.0 & 32.000 & 16.000 & 52.000 & 0.604 & 0.302 & 0.094 \\
fo 14.5 & 28.000 & 14.000 & 58.000 & 0.566 & 0.283 & 0.152 \\
fo 16.0 & 24.000 & 12.000 & 64.000 & 0.522 & 0.261 & 0.217 \\
q 1.5 & 0.000 & 85.000 & 15.000	& 0.000	& 0.958	& 0.042 \\
q 2.0 & 0.000 & 80.000 & 20.000 & 0.000 & 0.941 & 0.059 \\
q 2.5 & 0.000 & 75.000 & 25.000 & 0.000 & 0.923 & 0.077 \\
q 3.0 & 0.000 & 70.000 & 30.000 & 0.000 & 0.903 & 0.097 \\
q 3.5 & 0.000 & 65.000 & 35.000 & 0.000 & 0.881 & 0.119 \\
q 4.0 & 0.000 & 60.000 & 40.000 & 0.000 & 0.857 & 0.143 \\
q 4.5 & 0.000 & 55.000 & 45.000 & 0.000 & 0.830 & 0.170 \\
q 5.0 & 0.000 & 50.000 & 50.000 & 0.000 & 0.800 & 0.200 \\
br+q 2.25 & 25.000 & 50.000 & 25.000 & 25.000 & 50.000 & 0.000 \\
br+q 2.70 & 30.000 & 40.000 & 30.000 & 30.000 & 40.000 & 0.000 \\
br+q 3.2 & 35.556 & 28.889 & 35.556 & 35.556 & 28.889 & 0.000 \\
br+q 3.4 & 37.778 & 24.444 & 37.778 & 37.778 & 24.444 & 0.000
\end{tabular}
\end{table}

Capsules were created from 2 mm-diameter Pt$_{90}$Rh$_{10}$ and Au rods. They were cut by wire saw into 1 mm thick disks, into which were spark-eroded six holes in three rows, 250 microns wide and 700 microns deep. The capsules were cleaned by cycling between an acetone ultrasonic bath (15 minutes) and 1000 $^{\circ}$C furnace (20 minutes) three times. Any remaining contamination was removed with W$_{75}$Re$_{25}$ needle followed by a further trip to the ultrasonic bath. Capsule chambers were filled with powders of different compositions using a W$_{75}$Re$_{25}$ needle. Small pieces of tape were used to cover the other holes to avoid contamination.

Multianvil experiments were performed in the 5000 tonne press at the Bayerisches Geoinstitut (BGI). Cr-doped MgO octahedral multianvil assemblies with 18 mm edge length were used (Figure \ref{fig:assembly}). Two capsules were loaded into each assembly, with the open ends of the chambers facing each other and separated by six 0.05 mm thick Pt$_{90}$Rh$_{10}$ or Au foils. The assemblies were compressed to 13 GPa over four hours between eight tungsten carbide cubes with trucations of edge-length 11 mm. Pressure calibrations and details of the press can be found in \cite{FPTLDR2004} and \cite{KF2005}. The assemblies were then resistively heated via the stepped LaCrO$_3$ heater. The temperature was recorded using a W$_{97}$Re$_{3}$--W$_{75}$Re$_{25}$ (Type D) thermocouple inserted axially into the assembly. To avoid water loss, higher temperature runs were heated for a shorter duration. The experiments were quenched by cutting power to the heater. The assemblies were then decompressed to room pressure over 1000 minutes.

Capsules were recovered from the assembly, separated by wire saw and then ground by hand to reveal the tops of each capsule chamber. The wire saw was then used again to split the two sets of three chambers from each capsule. Each half-capsule was mounted in epoxy and mirror-polished. After revealing the edge of the capsule chambers, it was necessary to impregnate them with epoxy under vacuum before further polishing, in order to fill in the porous spaces created by the hydrous melt. Grinding under running water helped remove plucked grains from the polishing surface. Prepared and cleaned samples were then coated with a 10 nm thick carbon layer.

Analysis of run products was conducted via scanning electron microscope (SEM), using BSE imaging and EDS for phase identification (via the INCA software package).

\section{Results}
Experimental details and results are shown in Table \ref{table:experiments}. Superliquidus runs revealed networks of dendritic crystals of brucite, forsterite, enstatite and stishovite. Below the liquidus, solid phases tended to aggregate at the cold end of the capsules. Periclase formed equant crystals often associated with quench overgrowths which appear brighter in BSE images. Forsterite formed tabular crystals about 100 $\mu$m long. Enstatite crystals were much smaller, forming ca. 5 $\mu$m equant grains. Stishovite also formed equant to blocky grains, but crystals only separated from the melts in chambers with a high water content. In addition to the difference in texture and shape between quench crystals liquidus phases, subliquidus phases typically contained many small inclusions of Pt-Rh oxides. A typical set of run products is shown in Figure \ref{fig:sem}.

\begin{figure}[h!]
  \centering
      %\includegraphics[width=0.8\textwidth]{figures/experimental-ternary}
  \caption{The 18/11 octahedral assembly design used in this study.}
  \label{fig:assembly}
\end{figure}

\begin{landscape}
\begin{table}[ht!]
\caption{Experimental run conditions and run products determined by SEM/EDS. Compositions are listed in order of increasing molar H$_2$O content. Unused compositions for each run are marked by an en-dash. Minor solids are listed in brackets. Mineral abbreviations are as follows: br - brucite, per - periclase, co - hydroxychondrodite, en - clinoenstatite, s - stishovite.}
\label{table:experiments}
\begin{tabular}{llllccccc}
Expt \# & P (GPa) & T ($^{\circ}$C) & t (min) & brucite+PtAc & forsterite+PtAcid & enstatite+PtAcid & quartz+PtAcid & br+q \\
Z1063 & 13 & 1200 & 40 & -/br/br/br/L & co/fo/fo/(fo) & en/en/en/en & -/-/-/-/-/-/-/- & -/-/-/- \\
Z1085 & 13 & 1300 & 40 & -/-/per/L/L & -/fo/(en)/L & -/-/-/- & -/-/-/-/-/-/-/- & -/-/-/- \\
Z1079 & 13 & 1300 & 35 & -/per/per/L/L & fo/fo/en/(en) & en/en/en/en & -/-/-/-/-/-/-/- & -/-/-/- \\
Z1058 & 13 & 1400 & 30 & -/per/(per)/L/L & fo/fo/L/L & en/en/en/(en) & -/-/-/-/-/-/-/- & -/-/-/- \\
Z1060 & 13 & 1500 & 20 & -/-/-/-/- & fo/L/L/L & en/en/L/L & -/-/-/-/-/-/-/- & -/-/-/- \\
Z1140 & 13 & 1600 & 10 & -/-/-/-/- & -/-/-/- & -/-/-/- & s/s/s/s/s/s/s/(s) & s+en/en/en/en \\
Z1084 & 13 & 1650 & 10 & -/-/-/-/- & -/-/-/- & L/L/-/- & -/-/-/-/-/-/-/- & -/-/-/- \\
Z1207 & 13 & 1700 & 8 & -/-/-/-/- & -/-/-/- & -/-/-/- & s/s/s/s/s/s/(s)/L & s+en/(en)/L/L \\
Z1209 & 13 & 1800 & 5 & -/-/-/-/- & -/-/-/- & -/-/-/- & s/s/s/s/(s?)/L/L/L & s/L/L/L \\
Z1091 & 13 & 1900 & 5 & (per)/-/-/-/- & -/-/-/- & -/-/-/- & -/-/-/L/-/L/L/L & L/L/-/-
\end{tabular}
\end{table}
\end{landscape}

\begin{figure}[ht!]
  \centering
      \includegraphics[width=0.8\textwidth]{figures/Z1058_1400C_13GPa_fo}
  \caption{BSE image of typical run products. Experimental chambers from Experiment Z1058, conducted at 1400 $^{\circ}$C and 13 GPa. Chambers have compositions along the forsterite-water binary, with water contents increasing from left to right.}
  \label{fig:sem}
\end{figure}

\section{Discussion}
\subsection{Modelling water solubility in melts}

Oxygen in MgO-SiO$_2$-H$_2$O melts can exist either as molecular water, as a bridging oxygen between magnesium and/or silicon atoms, or as part of a terminal hydroxyl group. The general equation describing the reaction between melt species is
\begin{equation}
0.5 \textrm{H}_2\textrm{O} + 0.5 \textrm{O}_{br}^{2-} \rightleftharpoons \textrm{OH}_{tm}^-
\label{eqn:speciation}
\end{equation}

where the subscripts br and tm respectively refer to bridging and terminal oxygens. The equilibrium between these species can be described with an equilibrium constant $K$ \cite[e.g.][]{Stolper1982}
\begin{equation}
K = \frac{X_{\textrm{OH}_{tm}^-}}{\sqrt{X_{\textrm{H}_2\textrm{O}} X_{\textrm{O}_{br}^{2-}} } }
\label{eqn:equilibrium_constant}
\end{equation}

The equilibrium constant is related to the energy $\Delta G^{\circ}$ required to protonate a bridging oxygen
\begin{equation}
K = \exp\left(\frac{-\Delta G^{\circ}}{RT}\right)
\end{equation}

These equations can be used to describe in an average sense a range of melt-melt equilibria involving monomers (Mg(OH)$_2$ and Si(OH)$_4$), dimers (Mg$_2$O(OH)$_2$, MgSiO(OH)$_4$, Si$_2$O(OH)$_6$) and higher oligomers. A high proportion of monomers and dimers exist in relatively dilute solutions, but not in the concentrated solutions investigated in this study.

A number of models have been presented to investigate the behaviour of concentrated hydrous melts. \cite{SS1985} made the assumption that H$_2$O, O$_{br}^{2-}$ and OH$_{tm}^-$ mix ideally, with a parameter $r$ representing the number of oxygen atoms in each formula unit which are available for protonation. For example, an Mg$_2$SiO$_4$ melt with $r=4$ and $K=\infty$ represents a continuum between an anhydrous melt and one made of Mg(OH)$_2$ and Si(OH)$_4$ monomers. No distinction is made between isolated oligomers and partial depolymerisation of a silicate network; in other words, the energy required to form a terminal OH group is independent of local environment. Fixing $r$ places an implicit constraint on the maximum value of $K$ in water-rich compositions. For example, if $r=1$ in a hydrated MgO melt, then water contents exceeding 50 mol\% require that $K<\infty$. In practise, the equilibrium constant $K$ has been shown to be dependent on temperature and composition. \cite{SK1995} obtained a good fit to isochemical data with an expression of the form
\begin{equation}
\Delta G^{\circ} = a + b\,T
\end{equation}

\cite{THH2012} simplified the model of \cite{SS1985} by assuming that all hydrogen in the melt exists as OH$^-$, and that protonation is equally likely on any oxygen ($K=\infty$, $\Delta G^{\circ}=-\infty$, $r$ maximised). Deviations from this model must increase with increasing water content, and such a model cannot describe any melt more water-rich than the Mg(OH)$_2$-Si(OH)$_4$ binary.

To describe melting in the SiO$_2$-H$_2$O system, \cite{HM2012} generalised the aforementioned models by allowing non-ideality between the two endmembers. The excess term $\Delta G^{\circ}$(P,T) for the formation of OH$^-$ groups was assumed to be independent of composition; all oxygens are equally available for protonation, and the local extent of protonation does not affect the energy required for each further protonation.

In these models, equilibrium melt compositions are calculated by calculating the activity of a component $x$ in the melt which has the composition of the anhydrous liquidus phase $X$ from the Gibbs energy of melting of $X$ at the temperature of interest:
\begin{equation}
RT \ln a_{x,L}^X = G_{X_l} - G_{X_{s}} = - \int_T^{T_{melt}} S_{X_{l}}-S_{X_s} \, dT
\end{equation}
where $T_{melt}$ is the temperature of melting of the liquidus phase under anhydrous conditions.

In this study, we use a modified version of the molecular dynamics-derived anhydrous melt model of \cite{DKS2013} using melt endmembers MgO and SiO$_2$. For MgO melt, only the reference Helmholtz Free energy needed to be adjusted. For the SiO$_2$ melt, the bulk modulus of the SiO$_2$ endmember was increased to satisfy the experimental constraint that coexisting melt and coesite have similar volumes at 13--14 GPa. As a result, SiO$_2$ volumes at lower mantle conditions will be significantly higher than predicted, but this detail is not important for the present study. The volume at 1 bar is fit to experimental data. The entropy at reference conditions is adjusted by the value corresponding to the difference between the experimentally-derived and MD-calculated stishovite values. Remarkably, the value of the entropy derived in this way is essentially identical to that calculated for the experimental data of \cite{} near the coesite-stishovite-melt triple point. As such, we are confident that this value is accurate. It should be noted that the recent melting curve proposed by \cite{} (using a Simon's equation) predicts a much higher entropy of melting at 14 GPa. This is a result of downweighting the data in the study of \cite{}, which indicates a controversially high dT/dP. Nevertheless, this study is not inconsistent with higher pressure studies \citep{} if the gradient of the melting curve shallows between 30 and 70 GPa, which may be the consequence of changing Si-coordination and resultant densification in the melt.

A subregular solution model was adopted to describe the free energy between the MgO and SiO$_2$ melts. The two interaction parameters W$_{MS}$ and W$_{SM}$ 14 GPa were derived from the melt composition at the forsterite (fo)-HP clinoenstatite (hen) eutectic \citep{} and melting temperatures at the fo, hen and fo-hen cotectic compositions. The temperature dependence around 14 GPa was calculated by assuming an ideal excess entropy contribution, and then fitting to the known melting curves of forsterite and enstatite. The pressure dependencies were fit to the melting curves. It should be noted that although the excess volumes are similar to the results of \cite{DKS2013}, the excess entropies are some 50\% lower. However, they provide an excellent fit to the forsterite and orthoenstatite melting curves (see Supplementary Materials).

All of the calculations in the work were undertaken with the mineral physics toolkit Burnman \cite{}.

\citep{NM2008}
\citep{DM2010}
\citep{ZK1998}

\subsection{Subsystems}
\subsubsection{MgO-H$_2$O}
The dehydration of brucite at high pressure was previously studied by \cite{FIYKFO2005}. They reported that the stability of brucite reaches a maximum at about 9--10 GPa and 1200 $^{\circ}$C, decomposing at 1100--1150 $^{\circ}$C at 13 GPa. We observed large brucite crystals at 1200 $^{\circ}$C and 13 GPa; the temperature difference is probably within experimental uncertainties. In the present study, the composition of the fluid coexisting with brucite and periclase is constrained to be 64--66 mol\% H$_2$O. The MgO content of the fluid increases gradually with increasing pressure, as a result of the extremely high melting temperature of periclase. The extremely high melting point of MgO means that the best experimental estimate of the melting point at high pressure (7 GPa) is derived by extrapolation from iron-bearing (Mg,Fe)O, using the compositions of coexisting melt and solid to constrain the shape of the two-phase region \citep{ZF2008}. Their melting curve is within error of that obtained from early molecular dynamics simulations \citep{CG1994}, which suggest a melting temperature of $\sim$ 5373 K at 13 GPa. In this study, we follow more recent MD simulations \citep{KKS2013, DKS2013} which predict periclase melting at a lower temperature of $\sim$ 4180 K at 13 GPa, similar to that determined by previous ab-initio studies \citep{Alfe2005}, but significantly higher than early experimental estimates \citep[$\sim$ 3100 K][]{ZB1994}.

At 13 GPa, the calculated entropy of melting of periclase is $27.5$ J/K/mol, somewhat higher than the value from previous MD simulations \cite[$\sim$22 J/K/mol][]{CG1994}. This higher value will moderate the differences in melting temperature between the two studies. Using the ideal OH$^-$ mixing model of \cite{SS1985} with no compositional dependence on the formation of OH groups, $G^{\circ}= -75000$ provides a good fit to the data if additional asymmetry is added via a binary subregular model (W$_{MgO-H_2O}$=55000 J/mol, W$_{H_2O-MgO}$=0 J/mol). This asymmetry is required to fit the activity of water at the periclase-brucite-liquid triple point to the solid thermodynamic models of \citep{HP2011} and the equation of state of water formulated by \cite{PS1995} ($0.52$). This model also has the effect of increasing the tendency for melt-fluid immiscibility at the water-rich end of the system. The second critical endpoint of the MgO-H$_2$O is probably around 13 GPa \citep{MSUP2007}, so large positive interaction enthalpies are justified in this region of the phase diagram.

\begin{figure}[ht!]
  \centering
      \includegraphics[width=0.8\textwidth]{figures/per-H2O}
  \caption{The periclase-water phase diagram at 13 GPa, based on current experimental results. The position of the dry periclase melting point is taken from \cite{ZF2008}, with an entropy of dry melting from \cite{CG1994}. Model fits in grey use the model of \cite{SS1985}, with r=1 and K=$\infty$ (bottom), K=K(T) (expression in main text) and K=0 (all H2O as molecular, top). The black line includes an interaction term to match the fluid in equilibrium with brucite during it's decomposition to periclase. The black dotted line represents a regular solution model with MgO, Mg(OH)$_2$ and H$_2$O as distinct species in the melt (parameters in the main text).}
  \label{fig:MH}
\end{figure}

\clearpage
\subsubsection{Mg$_2$SiO$_4$-H$_2$O}
The melting point depression of forsterite in the presence of water at 13 GPa is illustrated in Figure \ref{fig:foH}. The anhydrous melting of forsterite is incongruent. The temperature of the fo $\rightarrow$ per + L reaction is taken from \citep{PW1993}, with the temperature of the metastable fo $\rightarrow$ L reaction extrapolated from the same study. The liquidus is calculated by interpolation between the melting temperature of periclase \citep{ZF2008} and an estimate of the temperature and composition of melt in equilibrium with periclase and forsterite by interpolating between the invariant fo-per-foL point \citep{PW1993} and position of the per liquidus at 16 GPa \citep{LF2012}.

One chamber from an experiment run at 1300 $^{\circ}$C revealed a single very small crystal of enstatite. Another run at 1200 $^{\circ}$C revealed large crystals of hydroxychondrodite in the chamber with the lowest water contents. Small differences in Mg:Si ratio could explain the differences in solid phases between the separate chambers, so it is assumed that the enstatite+forsterite and chondrodite+forsterite cotectics cross the forsterite+H$_2$O binary at high water contents and 1200--1300 $^{\circ}$C. This supposition is supported by \cite{DDFK2005}, who observe a double-crossing of the liquid and wadsleyite-water tie-line at 15 GPa, first at ca. 1100 $^{\circ}$C, and again at 1400 $^{\circ}$C. Very similar Mg:Si ratios were observed by \citep{LSOK2011} at 15 and 18.5 GPa.

\begin{figure}[ht!]
  \centering
      \includegraphics[width=0.8\textwidth]{figures/fo-H2O}
  \caption{The forsterite-water phase diagram at 13 GPa, based on current experimental results. The position of the dry melting point is taken from \citep{PW1993}. \citep{LF2012}.}
  \label{fig:foH}
\end{figure}
\clearpage
\subsubsection{MgSiO$_3$-H$_2$O}

The melting of clinoenstatite in the presence of water is shown in Figure \ref{fig:eoH}. The lack of stishovite as a liquidus phase is in agreement with \cite{YII2004}, who reported a clinoenstatite + stishovite + liquid field appearing at 13.5 GPa. 
\begin{figure}[ht!]
  \centering
      \includegraphics[width=0.8\textwidth]{figures/en-H2O}
  \caption{The clinoenstatite-water phase diagram at 13 GPa, based on current experimental results. The position of the dry melting point is taken from \cite{PG1990}.}
  \label{fig:eoH}
\end{figure}

\clearpage
\subsubsection{SiO$_2$-H$_2$O}
The 13 GPa SiO$_2$-H$_2$O binary phase diagram is shown in Figure \ref{fig:SH}. The anhydrous melting temperature of coesite and the temperature of the coesite-stishovite transition at 13 GPa are taken from \cite{ZLGHF1993}. The metastable melting temperature of stishovite is taken from the melting curve of \cite{Millotetal2015}, which is consistent with the shock data of \cite{LAM1983} and static melting data of \cite{ZLGHF1993}.

The liquidus in this system provides quite severe constraints on thermodynamics in the anhydrous system. Given that SiO$_2$-H$_2$O liquids approach ideality with increasing pressure and temperature \citep{HM2012}, we feel justified in using the ideal mixing model of \cite{SS1985}. Using this model constrains the entropy of melting of stishovite to be between about 8 and 19 J/K/mol, not significantly higher than room pressure values \citep{HM2012,ZLGHF1993}. A value of 13.5 J/K/mol with G$_{OH}$=35000-20T J/mol yields the model shown in Figure  \ref{fig:SH}. Unfortunately, these entropy values of are much higher than the values estimated from the melting curve of SiO$_2$. The melting curve of coesite has a negligable dT/dP at 13 GPa \citep{ZLGHF1993}, constraining the $\Delta$V of melting to be approximately zero. Using the volume of coesite at the coesite-stishovite-liquid triple junction as a proxy for the volume of the melt, published melting curves of stishovite \citep{Millotetal2015, ZLGHF1993} imply an entropy of melting of $\sim$120 J/K/mol. There are two potential resolutions to this problem. The first is that the assumptions of the hydrous melting model are inappropriate for the high pressure SiO$_2$ system. The second is that the thermodynamic properties in the anhydrous system need to be revisited.

%NB: Most likely culprit the change in S with temperature?? Need a melt and solid model to test this.

% We use the formulation of \cite{HM2012} to model SiO$_2$-H$_2$O melts at 13 GPa. The volume of stishovite is taken from \cite{HP2011}. The melting curve of coesite at this pressure has $dT/dP \sim 0$ \citep{ZLGHF1993}, so the volume of anhydrous SiO$_2$ at this pressure is roughly equal to the volume of coesite via the Clausium-Clapeyron relation $dT/dP = \Delta V / \Delta S$. Using these volumes, the metastable entropy of melting of stishovite $\Delta S_{stv-L}$ is estimated as 8.5 J/K/mol. For completeness, we also calculate the melting point and entropy of melting of ice VII at 13 GPa from \emph{P-V-T} data collected along the melting curve \citep{FFH2004}.

\begin{figure}[ht!]
  \centering
      \includegraphics[width=0.8\textwidth]{figures/SiO2-H2O}
  \caption{The silica-water phase diagram at 13 GPa, based on current experimental results. The liquidus curve is compared with previous experimentally-derived models computed at 1 GPa and 2 GPa \citep{HM2012}. An entropy of 13.5 J/K/mol is chosen to fit the data. Model fits use the model of \cite{SS1985}, with r=2 and K=$\infty$ (dotted grey), K=K(T) (black, expression in main text) and K=0 (dashed grey, all H2O as molecular).}
  \label{fig:SH}
\end{figure}

%\citep{SU2001}
%\citep{KOM2005}
%\citep{AA1980}
%\citep{Inoue1994}
%\citep{Wunder1998}
%\citep{AFRH2001}
%\citep{Nagano2002}
%\citep{MSUP2007}
%\citep{MFY2002}
%\citep{Sta2004}
%\citep{MFY2004}
%\citep{SUTG2001}

\clearpage
\subsection{The MgO-SiO$_2$-H$_2$O system}
The set of experimental run products can be used to create a preliminary liquidus diagram (Figure \ref{fig:liquidus}).

\cite{YII2004} suggested that there was a sharp break in the liquidus at the enstatite-stishovite cotectic (their Figure 6). However, this prediction breaks Schreinemakers Rules (specifically, that no one assemblage can occupy more than 180$^{\circ}$ around an invariant point).

\begin{figure}[ht!]
  \centering
      \includegraphics[width=0.8\textwidth]{figures/experimental-ternary}
  \caption{A preliminary liquidus diagram at 13 GPa, based on current experimental results.}
  \label{fig:liquidus}
\end{figure}

\section{Discussion}
\subsection{Water partitioning in the mantle transition zone}
% Olivine, wadsleyite and ringwoodite

\cite{DDFK2005}
\citep{LSOK2011}
\citep{OMY2000}

\subsection{Deep water release from subduction zones}
% Densities
\cite{MF2010}
\cite{BK2003}
\cite{MSK2008}

% Viscosities
% Shear viscosity 
% Density

%So at least in my models there are three quantities that are important for the pattern of melt migration, the permeability, the compaction viscosity and the shear viscosity (if you have elasticity/plasticity, other things come in, too). Everywhere in the equations where you have the permeability, it's divided by the melt viscosity, and the permeability itself is normally assumed to scale with phi^3 (or less common phi^2), and at least I didn't see any parametrisations describing the effect of temperature on the permeability, so I would assume it's small. So for the permeability, you can just use the values you get for your different P-T-X conditions, and see if the overall term permeability/melt_viscosity increases or decreases.

%For the viscosities of the solid you have the influence of temperature on the viscosity (which I think is similar for shear and compaction viscosity) and you have the effect of porosity, eta ~ exp(alpa * phi) with phi~-27 for the shear viscosity (I think this is what Rich Katz uses in his models, but I would have to look it up to be sure), and xi ~ phi0/phi for the compaction viscosity, and you can do the same here, compare how they change for your different P-T-X conditions.

\clearpage
\section*{References}

\bibliography{references_MSH}

\end{document}